\vspace{1.0cm} \Large \bf Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''

### 2016-09-22 mSystems/mSys_MAIN.tex
Some changes in main text and supplementary update leads to RC version
On branch master, changes to be committed:
	deleted:    doc/665380-revised-april.pdf
	new file:   doc/mSys-ITA-8-1-2016.pdf
	new file:   doc/mSys-sub-checklistAugust2016.pdf
	renamed:    mSys_suppl_0.1.pdf -> mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex
	new file:   mSys_suppl.pdf
	new file:   mSys_suppl.tex
	deleted:    mSys_suppl_0.1.tex
	modified:   option.tc
	new file:   results/finalPlot11.bkp.pdf
	modifie

% release codenumber (update it here!)
\newcommand\releasecode{0.5.0}

%%% PACKAGES
\usepackage{lmodern} % Activa fuentes y macros de Latin Modern
\usepackage[T1]{fontenc} % set font encoding
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage[english]{babel} 
\usepackage[expert, charter]{mathdesign} % Fuentes... disponibles conjuntos {charter|utopia}
%%\usepackage{calligra}
\usepackage[printwatermark]{xwatermark}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames,table]{xcolor}
\usepackage{parskip}
%%\usepackage{float}
%\usepackage[nolists,tablesfirst, nomarkers]{endfloat}
\usepackage{moreverb} % for verbatim output

@@

% watermarks (time consuming!)
\newwatermark[allpages,color=Fuchsia!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
\newwatermark[allpages,color=Fuchsia!12,angle=0,scale=1,xpos=0,ypos=137]{
\emph{mSystems} \LaTeX\ manuscript release \releasecode}

@@

\section*{Funding Information}
This work was supported by grants to AM from the Spanish Ministry of Science and Competitivity (projects SAF 2012-31187, SAF2013-49788-EXP, SAF2015-65878-R), Carlos III Institute of Health (projects PIE14/00045 and AC15/00022), Generalitat Valenciana (project PrometeoII/2014/065) and co-financed by FEDER, and grants to CPG from the Generalitat Valenciana Prometeo Grants II/2014/050, II/2014/065, 419 by the Spanish Grants FPA2011-29678, BFU2012-39816-C02-01 of MINECO and by PITN-GA-420 2011-289442-INVISIBLES. From the Ministry of Economy and Competitiveness (grant FPI BES-2013-062767).

 \todo{Modificar becas de JMM y DMM ¿poner algún grant más?}

@@

@@

@@

@@

@@

@@

% main studies
\bibitem{moving} {\bf Caporaso JG, Lauber CL, Costello EK, Berg-Lyons D, Gonzalez A, Stombaugh J, Knights D, Gajer P, Ravel J, Fierer N, Gordon JI, Knight R.} 2011. Moving pictures of the human microbiome. Genome Biol {\bf 12}:R50.
\bibitem{LEA} {\bf Faith JJ, Guruge JL, Charbonneau M, Subramanian S, Seedorf H, Goodman AL, Clemente JC, Knight R, Heath AC, Leibel RL, Rosenbaum M, Gordon JI.} 2013. The long-term stability of the human gut microbiota. Science (80- ) {\bf 341}:1237439.
\bibitem{kwashiorkor} {\bf Smith MI, Yatsunenko T, Manary MJ, Trehan I, Mkakosya R, Cheng J, Kau AL, Rich SS, Concannon P, Mychaleckyj JC, Liu J, Houpt E, Li J V, Holmes E, Nicholson J, Knights D, Ursell LK, Knight R, Gordon JI.} 2013. Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. Science {\bf 339}:548–54.
\bibitem{diet} {\bf David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe BE, Ling A V, Devlin AS, Varma Y, Fischbach MA, Biddinger SB, Dutton RJ, Turnbaugh PJ.} 2014. Diet rapidly and reproducibly alters the human gut microbiome. Nature {\bf 505}:559–63.
\bibitem{antibiotic} {\bf Dethlefsen L, Relman DA.} 2011. Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation. Proc Natl Acad Sci {\bf108}:4554–61.

\bibitem{FASTX} {\bf Gordon, A., Hannon, G.J.} 2010. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools v0.0.13. http://hannonlab.cshl.edu/fastx\_toolkit/ (last accessed 26 Jul 2016).
\bibitem{QIIME} {\bf Caporaso JG, Kuczynski J, Stombaugh J, Bittinger K, Bushman FD, Costello EK, Fierer N, Peña AG, Goodrich JK, Gordon JI, Huttley G a, Kelley ST, Knights D, Koenig JE, Ley RE, Lozupone C a, Mcdonald D, Muegge BD, Pirrung M, Reeder J, Sevinsky JR, Turnbaugh PJ, Walters W a, Widmann J, Yatsunenko T, Zaneveld J, Knight R.} 2010. correspondence QIIME allows analysis of high- throughput community sequencing data Intensity normalization improves color calling in SOLiD sequencing. Nat Publ Gr {\bf 7}:335–336.
\bibitem{SILVA} {\bf Quast C, Pruesse E, Yilmaz P, Gerken J, Schweer T, Yarza P, Peplies J, Glöckner FO}. 2013. The SILVA ribosomal RNA gene database project: improved data processing and web-based tools. Acids Res. {\bf 41}:D590-D596
\bibitem{LMAT} {\bf Ames SK, Hysom DA, Gardner SN, Lloyd GS, Gokhale MB, Allen JE}.  2013. Scalable metagenomic taxonomy classification using a reference genome database.  Bioinformatics {\bf 29}:2253-2260.
\bibitem{LMAT2} {\bf Ames SK, Gardner SN, Marti JM, Slezak TR, Gokhale MB, Allen JE}. 2015. Using populations of human and microbial genomes for organism detection in metagenomes. Genome Res. {\bf 25}:1056-67.
\bibitem{fs} {\bf Eisler Z, Bartos I, Kertész J.} 2008. Fluctuation scaling in complex systems: Taylor’s law and beyond1. Adv Phys {\bf 57}:89–142.
\bibitem{stat} {\bf Jørgensen B, Martinez JR, Tsao M.} 1994. Asymptotic behaviour of the variance function. Scand J Stat {\bf 21}:223–243.

\bibitem{convergence1} {\bf Fronczak A, Fronczak P.} 2010. Origins of Taylor’s power law for fluctuation scaling in complex systems. Phys Rev E {\bf 81}:066112.
\bibitem{convergence2} {\bf Kendal, W.S., Jorgensen,B.} Taylor's power law and fluctuation scaling explained by a central-limit-like convergence.  Phys. Rev. E {\bf 83} 066115.
\bibitem{convergence3} {\bf Kendal WS, Jørgensen B.} 2011. Tweedie convergence: A mathematical basis for Taylor’s power law, 1/f noise, and multifractality. Phys Rev E {\bf 84}:066120.

%discussion
\bibitem {kilpatrick} {\bf Kilpatrick a M, Ives a R.} 2003. Species interactions can explain Taylor’s power law for ecological time series. Nature {\bf 422}:65–68.
\bibitem {ballantyne} {\bf Ballantyne IV F, J. Kerkhoff A.} 2007. The observed range for temporal mean-variance scaling exponents can be explained by reproductive correlation. Oikos {\bf 116}:174–180.
\bibitem {joao} {\bf Stein RR, Bucci V, Toussaint NC, Buffie CG, Rätsch G, Pamer EG, Sander C, Xavier JB.}  2013. Ecological modeling from time-series inference: insight into dynamics and stability of intestinal microbiota. PLoS Comput Biol {\bf 9}:e1003388.
\bibitem {mehta} {\bf Fisher CK, Mehta P.} 2014. Identifying keystone species in the human gut microbiome from metagenomic timeseries using sparse linear regression. PLoS One {\bf 9}:e102451.
\bibitem {bucci}  {\bf Bucci V, Tzen B, Li N, Simmons M, Tanoue T, Bogart E, Deng L, Yeliseyev V, Delaney ML, Liu Q, Olle B, Stein RR, Honda K, Bry L, Gerber GK.} 2016. MDSINE: Microbial Dynamical Systems INference Engine for microbiome time-series analyses. Genome Biol {\bf 17}:121.
\bibitem {cohen_taylor} {\bf Cohen JE, Xu M, Schuster WSF}. 2013. Stochastic multiplicative population growth predicts and interprets Taylor ’ s power law of fluctuation scaling Stochastic multiplicative population growth predicts and interprets Taylor ’ s power law of fluctuation scaling. Proc R Soc B Biol Sci {\bf 280}:20122955.
\bibitem {koenig} {\bf Koenig JE, Spor A, Scalfone N, Fricker AD, Stombaugh J, Knight R, Angenent LT, Ley RE.} 2011. Succession of microbial consortia in the developing infant gut microbiome. Proc Natl Acad Sci {\bf 108}:4578–4585.
\bibitem {FD} {\bf Weber, J.} 1956. Fluctuation Dissipation Theorem. Phys. Rev. {\bf 101}:1620-6
\bibitem {tikhonov} {\bf  Tikhonov M.} 2016. Community-level cohesion without cooperation. Elife {\bf 5}.

%M&M
\bibitem{biopwrlaw} {\bf Xiao X, White EP, Hooten MB, Durham SL}. 2011. On the use of log-transformation vs. nonlinear regression for analyzing biological power laws. Ecology {\bf 92}:1887-1894
\bibitem{boot} {\bf Wu, C.F.J.} 1986. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}:1261-1350 
\end{thebibliography}

\end{document}

>>>>

% release codenumber (update it here!)
\newcommand\releasecode{$\beta$}

%%% PACKAGES
\usepackage{lmodern} % Activa fuentes y macros de Latin Modern
\usepackage[T1]{fontenc} % set font encoding
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage[english]{babel} 
\usepackage[expert, charter]{mathdesign} % Fuentes... disponibles conjuntos {charter|utopia}
%%\usepackage{calligra}
\usepackage[printwatermark]{xwatermark}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames,table]{xcolor}
\usepackage{parskip}
%%\usepackage{float}
\usepackage[nolists,tablesfirst, nomarkers]{endfloat}
\usepackage{moreverb} % for verbatim output

@@

% watermarks (time consuming!)
\newwatermark[allpages,color=Green!4,angle=45,scale=12,xpos=-60,ypos=0]{RC\releasecode}
\newwatermark[allpages,color=Green!12,angle=0,scale=1,xpos=0,ypos=137]{
\emph{mSystems} \LaTeX\ manuscript RC\releasecode}

@@

\section*{Funding Information}
This work was supported by grants to AM from the Spanish Ministry of Science and Competitivity (projects SAF2012-31187, SAF2013-49788-EXP, SAF2015-65878-R), Carlos III Institute of Health (projects PIE14/00045 and AC15/00022), Generalitat Valenciana (project PrometeoII/2014/065) and co-financed by FEDER, and grants to CPG from the Generalitat Valenciana Prometeo Grants II/2014/050, II/2014/065, 419 by the Spanish Grants FPA2011-29678, BFU2012-39816-C02-01 of MINECO and by PITN-GA-420 2011-289442-INVISIBLES. From the Ministry of Economy and Competitiveness (grants FPI BES-2012-052900 and FPI BES-2013-062767).

%\clearpage

@@

%% Introduction

@@

\bibitem{mind} {\bf Cryan JF, Dinan TG}. 2012. Mind-altering microorganisms: the impact of the gut microbiota on brain and behaviour. Nat Rev Neurosci {\bf 13}:701–712
\bibitem{AD} {\bf Xu R, Wang Q}. 2016. Towards understanding brain-gut-microbiome connections in Alzheimer's disease. BMC Systems Biology {\bf 10}:277-285

@@

%

@@

\bibitem{panthropology} {\bf Amato KR} 2016. Incorporating the Gut Microbiota Into Models of Human and Non-Human Primate Ecology and Evolution. Yearbook Of Physical Anthropology {\bf 157}:S196–S215.

@@

\bibitem {cohen_taylor} {\bf Cohen JE, Xu M, Schuster WSF}. 2013. Stochastic multiplicative population growth predicts and interprets Taylor’s power law of fluctuation scaling. Proc R Soc B Biol Sci {\bf 280}:20122955.

@@

%% Results
\bibitem{fs} {\bf Eisler Z, Bartos I, Kertész J.} 2008. Fluctuation scaling in complex systems: Taylor’s law and beyond1. Adv Phys {\bf 57}:89–142.
% -> main studies
\bibitem{moving} {\bf Caporaso JG, Lauber CL, Costello EK, Berg-Lyons D, Gonzalez A, Stombaugh J, Knights D, Gajer P, Ravel J, Fierer N, Gordon JI, Knight R.} 2011. Moving pictures of the human microbiome. Genome Biol {\bf 12}:R50.
\bibitem{antibiotic} {\bf Dethlefsen L, Relman DA.} 2011. Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation. Proc Natl Acad Sci {\bf108}:4554–61.
\bibitem{LEA} {\bf Faith JJ, Guruge JL, Charbonneau M, Subramanian S, Seedorf H, Goodman AL, Clemente JC, Knight R, Heath AC, Leibel RL, Rosenbaum M, Gordon JI.} 2013. The long-term stability of the human gut microbiota. Science {\bf 341}:1237439.
\bibitem{kwashiorkor} {\bf Smith MI, Yatsunenko T, Manary MJ, Trehan I, Mkakosya R, Cheng J, Kau AL, Rich SS, Concannon P, Mychaleckyj JC, Liu J, Houpt E, Li J V, Holmes E, Nicholson J, Knights D, Ursell LK, Knight R, Gordon JI.} 2013. Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. Science {\bf 339}:548–54.
\bibitem{diet} {\bf David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe BE, Ling A V, Devlin AS, Varma Y, Fischbach MA, Biddinger SB, Dutton RJ, Turnbaugh PJ.} 2014. Diet rapidly and reproducibly alters the human gut microbiome. Nature {\bf 505}:559–63.
% <- END main studies
\bibitem{stat} {\bf Jørgensen B, Martinez JR, Tsao M.} 1994. Asymptotic behaviour of the variance function. Scand J Stat {\bf 21}:223–243.
\bibitem{convergence1} {\bf Fronczak A, Fronczak P.} 2010. Origins of Taylor’s power law for fluctuation scaling in complex systems. Phys Rev E {\bf 81}:066112.
\bibitem{convergence2} {\bf Kendal, W.S., Jorgensen,B.} Taylor's power law and fluctuation scaling explained by a central-limit-like convergence.  Phys. Rev. E {\bf 83} 066115.
\bibitem{convergence3} {\bf Kendal WS, Jørgensen B.} 2011. Tweedie convergence: A mathematical basis for Taylor’s power law, 1/f noise, and multifractality. Phys Rev E {\bf 84}:066120.

%% Discussion
\bibitem {kilpatrick} {\bf Kilpatrick a M, Ives a R.} 2003. Species interactions can explain Taylor’s power law for ecological time series. Nature {\bf 422}:65–68.
\bibitem {ballantyne} {\bf Ballantyne IV F, J. Kerkhoff A.} 2007. The observed range for temporal mean-variance scaling exponents can be explained by reproductive correlation. Oikos {\bf 116}:174–180.
\bibitem {joao} {\bf Stein RR, Bucci V, Toussaint NC, Buffie CG, Rätsch G, Pamer EG, Sander C, Xavier JB.}  2013. Ecological modeling from time-series inference: insight into dynamics and stability of intestinal microbiota. PLoS Comput Biol {\bf 9}:e1003388.
\bibitem {mehta} {\bf Fisher CK, Mehta P.} 2014. Identifying keystone species in the human gut microbiome from metagenomic timeseries using sparse linear regression. PLoS One {\bf 9}:e102451.
\bibitem {bucci}  {\bf Bucci V, Tzen B, Li N, Simmons M, Tanoue T, Bogart E, Deng L, Yeliseyev V, Delaney ML, Liu Q, Olle B, Stein RR, Honda K, Bry L, Gerber GK.} 2016. MDSINE: Microbial Dynamical Systems INference Engine for microbiome time-series analyses. Genome Biol {\bf 17}:121.
\bibitem {koenig} {\bf Koenig JE, Spor A, Scalfone N, Fricker AD, Stombaugh J, Knight R, Angenent LT, Ley RE.} 2011. Succession of microbial consortia in the developing infant gut microbiome. Proc Natl Acad Sci {\bf 108}:4578–4585.
\bibitem {FD} {\bf Weber, J.} 1956. Fluctuation Dissipation Theorem. Phys. Rev. {\bf 101}:1620-6
\bibitem {tikhonov} {\bf  Tikhonov M.} 2016. Community-level cohesion without cooperation. Elife {\bf 5}.

%% M&M
% Selection and Methods
\bibitem{Einstein} {\bf Einstein A.} 1905. Über die von der molekularkinetischen Theorie der Wärme geforderte Bewegung von in ruhenden Flüssigkeiten suspendierten Teilchen. Annalen der Physik, {\bf 322}:549-560.
\bibitem{QIIME} {\bf Caporaso JG, Kuczynski J, Stombaugh J, Bittinger K, Bushman FD, Costello EK, Fierer N, Peña AG, Goodrich JK, Gordon JI, Huttley G a, Kelley ST, Knights D, Koenig JE, Ley RE, Lozupone C a, Mcdonald D, Muegge BD, Pirrung M, Reeder J, Sevinsky JR, Turnbaugh PJ, Walters W a, Widmann J, Yatsunenko T, Zaneveld J, Knight R.} 2010. correspondence QIIME allows analysis of high- throughput community sequencing data Intensity normalization improves color calling in SOLiD sequencing. Nat Publ Gr {\bf 7}:335–336.
\bibitem{LMAT} {\bf Ames SK, Hysom DA, Gardner SN, Lloyd GS, Gokhale MB, Allen JE}.  2013. Scalable metagenomic taxonomy classification using a reference genome database.  Bioinformatics {\bf 29}:2253-2260.
\bibitem{FASTX} {\bf Gordon, A, Hannon, GJ.} 2010. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools v0.0.13. http://hannonlab.cshl.edu/fastx\_toolkit/ (last accessed 26 Jul 2016).
\bibitem{SILVA} {\bf Quast C, Pruesse E, Yilmaz P, Gerken J, Schweer T, Yarza P, Peplies J, Glöckner FO}. 2013. The SILVA ribosomal RNA gene database project: improved data processing and web-based tools. Acids Res. {\bf 41}:D590-D596
\bibitem{LMAT2} {\bf Ames SK, Gardner SN, Marti JM, Slezak TR, Gokhale MB, Allen JE}. 2015. Using populations of human and microbial genomes for organism detection in metagenomes. Genome Res. {\bf 25}:1056-67.
\bibitem{boot} {\bf Wu, C.F.J.} 1986. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}:1261-1350 
\bibitem{biopwrlaw} {\bf Xiao X, White EP, Hooten MB, Durham SL}. 2011. On the use of log-transformation vs. nonlinear regression for analyzing biological power laws. Ecology {\bf 92}:1887-1894
\end{thebibliography}
\end{document}

### 2016-09-22 mSystems/mSys_MandM.tex
Some changes in main text and supplementary update leads to RC version
On branch master, changes to be committed:
	deleted:    doc/665380-revised-april.pdf
	new file:   doc/mSys-ITA-8-1-2016.pdf
	new file:   doc/mSys-sub-checklistAugust2016.pdf
	renamed:    mSys_suppl_0.1.pdf -> mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex
	new file:   mSys_suppl.pdf
	new file:   mSys_suppl.tex
	deleted:    mSys_suppl_0.1.tex
	modified:   option.tc
	new file:   results/finalPlot11.bkp.pdf
	modifie

%%
\subsection*{Model} \label{sec:model}

We model the microbial abundances across time along the lines of Blumm \textit{et al.}\cite{ranking}. The dynamics of taxon relative abundances is described by the Langevin equation:
\begin{linenomath}
\begin{equation*}
\dot{x_i} = F_i \cdot x_i^\alpha + V \cdot x_i^\beta \xi_i(t) - \phi(t) \cdot x_i,
\end{equation*}
\end{linenomath}
where F$_i$ captures the fitness of the taxon i, V corresponds to the noise amplitude and $\xi_i$(t) is a Gaussian random noise with zero mean  $<\xi_i(t)>$ = 0 and variance uncorrelated in time, $<\xi_i(t) \xi_i(t')>$ =  $\delta(t'-t)$, . The function $\phi(t)$ ensures the normalization at all times, $\sum x_i(t) = 1$, and corresponds to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$.
The temporal evolution of the probability that a taxon i has a relative abundance $x_i(t)$, P(x$_i$,t), is determined by the Fokker-Planck equation:
\begin{linenomath}
\begin{equation*}
\frac{\partial P}{\partial t} = - \frac{\partial}{\partial x_i}  [(F_i \cdot x_i^\alpha - \phi(t) \cdot x_i ) \cdot P]+ \frac{1}{2} \frac{\partial^2}{\partial x_i^2} (V^2 \cdot x_i^{2\beta}\cdot P).
\end{equation*}
\end{linenomath}
The microbiota evolves towards a steady-state with a time-independent probability depending on the values of $\alpha$, $\beta$, F$_i$ and V. For $\alpha<1$ (otherwise, systems are always unstable), the steady-state probability may be localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness F$_i$  dominates or is overwhelmed by the noise amplitude V. The steady-state solution of the Fokker-Planck equation is given by:
\begin{eqnarray}
P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \cdot x_i^{-2\beta}  \cdot \exp\Big[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \notag \\
P_0 (x_i) &=& C_e(\alpha,\beta,F_i,V)  \cdot x_i^{\frac{2F_i}{V^2} -2\beta}  \cdot \exp\Big[\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta = 1+\alpha,
\notag
\end{eqnarray}
where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and C$_{ne}$ and C$_{e}$ are integrals that should be solved numerically for the parameters of interest. The ordered phase happens when the solution has a maximum in the physical interval ($0<x_i<1$). For larger V, the transition to a disordered phase happens when the maximum shifts to the unphysical region $x_i<0$, which sets the phase transition region V($\alpha,\beta,F_i$). The phase transition region can be calculated analytically in particular cases:
\begin{eqnarray}
F_i^2 &=& 4 \beta \phi_0 V^2\quad \textrm{if} \quad  \beta = \alpha \neq 1, \notag \\
F_i &=& \beta V^2\quad \textrm{if} \quad  2\beta = 1+\alpha,
\notag
\end{eqnarray}
where the first case, simplifies to $F= 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominates in $\phi_0$. 
In many physical systems (Brownian motion is the classical example), the two terms of the Langevin equation are related.  The \emph{Fluctuation Dissipation Theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}. The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations 
of the microbiota, assuming that it is in equilibrium (the ordered phase).

\todo{Explain better the Fluctuation Dissipation Theorem}

@@

The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{QIIME}. WGS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold. Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level. We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure \ref{fig:taxlev1}).

@@

Metadata of each study is provided in Supplementary Tables S1 to S4. All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, where both shotgun metagenomic sequencing (SMS) and 16S rRNA were used. In the latter case we selected to work with SMS data to show that our method is valid regardless of the source of taxonomic information. Each one of the datasets was treated as follows:

@@

%%
\subsubsection*{Taxa level selection}
We selected genus as taxonomic level for the subsequent steps of our work. In order to ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see last subsection of Material and Methods), we tested two different data sets. In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels. The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels. The Figures \ref{fig:taxlev1} (overview) and \ref{fig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.

\begin{figure} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
\caption{Overview of comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space. For 16S (former row of subfigures), the levels are family vs. genus, whereas for SMS (latter row of subfigures) levels are genus vs. species. The left column shows the raw results and the right column plots the standardized results (see Section \ref{sec:stan})}
\label{fig:taxlev1}
\end{figure}

@@

>>>>

%%
%\small
\subsection*{Model} \label{sec:model}

We model the microbial abundances across time along the lines of Blumm \textit{et al.}\cite{ranking}. The dynamics of taxon relative abundances is described by the Langevin equation:
\begin{linenomath}
\begin{equation*}
\dot{x_i} = F_i \cdot x_i^\alpha + V \cdot x_i^\beta \xi_i(t) - \phi(t) \cdot x_i,
\end{equation*}
\end{linenomath}
where F$_i$ captures the fitness of the taxon i, V corresponds to the noise amplitude and $\xi_i$(t) is a Gaussian random noise with zero mean  $<\xi_i(t)>$ = 0 and variance uncorrelated in time, $<\xi_i(t) \xi_i(t')>$ =  $\delta(t'-t)$, . The function $\phi(t)$ ensures the normalization at all times, $\sum x_i(t) = 1$, and corresponds to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$.
The temporal evolution of the probability that a taxon i has a relative abundance $x_i(t)$, P(x$_i$,t), is determined by the Fokker-Planck equation:
\begin{linenomath}
\begin{equation*}
\frac{\partial P}{\partial t} = - \frac{\partial}{\partial x_i}  [(F_i \cdot x_i^\alpha - \phi(t) \cdot x_i ) \cdot P]+ \frac{1}{2} \frac{\partial^2}{\partial x_i^2} (V^2 \cdot x_i^{2\beta}\cdot P).
\end{equation*}
\end{linenomath}
The microbiota evolves towards a steady-state with a time-independent probability depending on the values of $\alpha$, $\beta$, F$_i$ and V. For $\alpha<1$ (otherwise, systems are always unstable), the steady-state probability may be localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness F$_i$  dominates or is overwhelmed by the noise amplitude V. The steady-state solution of the Fokker-Planck equation is given by:
\begin{eqnarray}
P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \cdot x_i^{-2\beta}  \cdot \exp\Big[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \notag \\
P_0 (x_i) &=& C_e(\alpha,\beta,F_i,V)  \cdot x_i^{\frac{2F_i}{V^2} -2\beta}  \cdot \exp\Big[\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta = 1+\alpha,
\notag
\end{eqnarray}
where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and C$_{ne}$ and C$_{e}$ are integrals that should be solved numerically for the parameters of interest. The ordered phase happens when the solution has a maximum in the physical interval ($0<x_i<1$). For larger V, the transition to a disordered phase happens when the maximum shifts to the unphysical region $x_i<0$, which sets the phase transition region V($\alpha,\beta,F_i$). The phase transition region can be calculated analytically in particular cases:
\begin{eqnarray}
F_i^2 &=& 4 \beta \phi_0 V^2\quad \textrm{if} \quad  \beta = \alpha \neq 1, \notag \\
F_i &=& \beta V^2\quad \textrm{if} \quad  2\beta = 1+\alpha,
\notag
\end{eqnarray}
where the first case, simplifies to $F= 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominates in $\phi_0$. 
In many physical systems (Brownian motion is the classical example \cite{Einstein}), the two terms of the Langevin equation are related.  The \emph{Fluctuation Dissipation Theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}. The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations 
of the microbiota, assuming that it is in equilibrium (the ordered phase).

@@

The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{QIIME}. Shotgun metagenomic sequencing (SMS) data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold. Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level. We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure \ref{fig:taxlev1}).

@@

Metadata of each study is provided in Supplementary Tables S1 to S4. All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, where both SMS and 16S rRNA data were used. In the latter case we selected to work with SMS data to show that our method is valid regardless of the source of taxonomic information. Each one of the datasets was treated as follows:

@@

%%
\subsubsection*{Taxa level robustness}
We selected genus as taxonomic level for the subsequent steps of our work. In order to ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see last subsection of Material and Methods), we tested two different data sets. In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels. The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels. The Figures \ref{fig:taxlev1} (overview) and \ref{fig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.

\begin{figure} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
\caption{Overview of comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space. The former row of subfigures is for 16S, where levels are family (blue circles) vs. genus (purple triangles). Whereas the latter row of subfigures is for SMS, where levels are genus (blue circles) vs. species (purple triangles). The left column shows the raw results and the right column plots the standardized results (see Standardization under Material and Methods).}
\label{fig:taxlev1}
\end{figure}

@@

%\normalsize
